---
title: "Prueba"
author: "María Ponce"
date: "2023-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, results = 'hide',
                      warning = FALSE, cache = TRUE)
```


```{r}
############# LIBRARIES #############

library(readxl)     # Load file
library(survival)   # Surivival Analysis
library(dplyr)      # Process data
library(fabricatr)  # Divide data into quartiles

library(ggplot2)    # Plots
library(survminer)  # Survival plots
library(ggthemes)   # Themes
```


## Introduction

We use the tool named Cancertool. Using this tool we performed a basic statistical analysis using the Glinsky, Grasso, Taylor and TCGA databases. The analysis we wanted to performed was focused on the Gleason score of the genes PTEN, SOX9 and PML in the primary tumor group (PT). 

*Grasso database does not include Gleason score.*

The main objective of this file is to understand the basis of the DFS analysis which is performed on a regular basis in the Lab, to accomplish it we focused on the Glinsky database.

The following websites were used: 

> <https://cran.r-project.org/web/packages/survival/survival.pdf>
> <https://rviews.rstudio.com/2017/09/25/survival-analysis-with-r/>
> <https://bioconnector.github.io/workshops/r-survival.html#kaplan-meier_plots>
> <https://www.r-bloggers.com/2021/08/log-rank-test-in-r-survival-curve-comparison/>
> <https://whitlockschluter3e.zoology.ubc.ca/RExamples/Rcode_Chapter_21.html#hazard_ratio>
> <https://rpkgs.datanovia.com/survminer/#installation-and-loading>

```{r }
############# LOAD DATASETS #############

## Select Glinksy database per gene (sheet = 3)
data_pml <- read_xlsx("1679306381_PCa_BasicAnalysis\\DiseaseFreeSurvival\\PML\\DFSRAWdata.xlsx", sheet = 3)
# data_pten <- read_xlsx("1679306381_PCa_BasicAnalysis\\DiseaseFreeSurvival\\PTEN\\DFSRAWdata.xlsx", sheet = 3)
# data_sox9 <- read_xlsx("1679306381_PCa_BasicAnalysis\\DiseaseFreeSurvival\\SOX9\\DFSRAWdata.xlsx", sheet = 3)

# Verify
colnames(data_pml)
head(data_pml)

############# PREPROCESSING #############

# Recurrence status as numeric 1=0;censored and 2=1;event 
data_pml$`Recurrence status` <- as.numeric(data_pml$`Recurrence status`)

# Summary
str(data_pml)

# Sort data by time to recurrence
data_pml %>% arrange(`Time to recurrence (months)`)

############# DFS CURVE QUANTILES #############

# Simplify version of the analysis 
# based on the data found in the file

### Kaplan - Meier survival curve ### 

## Survival Object
km_curve <- with(data_pml, Surv(`Time to recurrence (months)`,`Recurrence status`))
print(km_curve)

## Fit survival curve based on the Quartile variable
km_fit <- survfit(Surv(`Time to recurrence (months)`,`Recurrence status`) ~ Quartile, data = data_pml)
summary(km_fit)

## Survival curves basic plot 
plot(km_fit)

### Log Rank test ratio ### 
logrank <- survdiff(Surv(`Time to recurrence (months)`,`Recurrence status`) ~ Quartile, data = data_pml)
## Summary log rank test
print(logrank)
## p-value
print(logrank$pvalue)

### Hazard ratio (Q1 vs Q4)### 
HR <- (logrank$obs[1]/logrank$obs[4])/(logrank$exp[1]/logrank$exp[4])
print(HR)
## Proportional hazard regression model 
# Log Rank ratio and HR can be obtained using this function
HR_lab <- coxph(km_curve ~ Quartile,data = data_pml)
summary(HR_lab)
```


```{r}
############# ADVANCED PREPROCESSING #############
data <- data_pml %>% mutate(data_pml,
                        # Quantile variable
                        Quantiles_div = split_quantile(`Log2 mRNA expression`,type = 4), # type = 4 = quantiles
                        # Median variable
                        Median_div = split_quantile(`Log2 mRNA expression`,type = 2) # type = 2 = median
                ) 
## Survival Object
km_curve <- with(data, Surv(`Time to recurrence (months)`,`Recurrence status`))
print(km_curve)

## Fit survival curve based on the Quartile variable
km_fit <- survfit(Surv(`Time to recurrence (months)`,`Recurrence status`) ~ Median_div, data = data)
summary(km_fit)


############# ADVANCED GRAPHICS #############
ggsurvplot(km_fit, data = data, 
            pval = TRUE,
            ggtheme = theme_few()
           )
```



```{r}
# data_glinsky <- readxl::read_xlsx("CompleteRAWdata.xlsx", sheet = 2)
# head(data_glinsky)
# summary(data_glinsky)
# 
# data_grasso <- readxl::read_xlsx("CompleteRAWdata.xlsx", sheet = 3)
# head(data_grasso)
# summary(data_grasso)
# 
# data_taylor <- readxl::read_xlsx("CompleteRAWdata.xlsx", sheet = 4)
# head(data_taylor)
# summary(data_taylor)
# 
# data_tcga <- readxl::read_xlsx("CompleteRAWdata.xlsx", sheet = 5)
# head(data_tcga)
# summary(data_tcga)
```
